### 2019-02-22 report/final-report.tex
experiments

\paragraph*{Benchmark Suite Construction}
We constructed our benchmarks primarily by adapting examples from
Augeas~\cite{augeas} and 
Flash Fill~\cite{gulwani-popl-2014}.

Augeas is a configuration editing system for Linux that uses lens
combinators similar to those in Boomerang. However, it transforms
strings on the left to structured trees on the right rather than
transforming strings to strings.
We adapted these Augeas lenses to our setting by converting the
right-hand sides to strings that correspond to serialized versions
of the tree formats.  
We derived 29 of the benchmark tests by
adapting the first 27 lenses in alphabetical order, as well as the lenses
\CF{aug/xml-firstlevel} and \CF{aug/xml} that were referenced
by the `A' lenses.
Furthermore, the 12 last synthesis problems derived
from Augeas were tested after \Optician{} was
finalized, demonstrating that the optimizations were not
overtuned to perform well on the testing data.

Flash Fill is a system that allows users to specify common string
transformations by example~\cite{gulwani-popl-2014}.  
We derived three benchmarks from the first few examples in the
paper and one from the running example on
extracting phone numbers.

@@

>>>>

\paragraph*{Benchmark Suite Construction}
We constructed our benchmarks primarily by adapting examples from
Augeas~\cite{augeas} and 
Flash Fill~\cite{gulwani-popl-2014}.
%
Augeas is a configuration editing system for Linux that uses lens
combinators similar to those in Boomerang. However, it transforms
strings on the left to structured trees on the right rather than
transforming strings to strings.
We adapted these Augeas lenses to our setting by converting the
right-hand sides to strings that correspond to serialized versions
of the tree formats.  
We derived 29 of the benchmark tests by
adapting the first 27 lenses in alphabetical order, as well as the lenses
\CF{aug/xml-firstlevel} and \CF{aug/xml} that were referenced
by the `A' lenses.
Furthermore, the 12 last synthesis problems derived
from Augeas were tested after \Optician{} was
finalized, demonstrating that the optimizations were not
overtuned to perform well on the testing data.
%
Flash Fill is a system that allows users to specify common string
transformations by example~\cite{gulwani-popl-2014}.  
We derived three benchmarks from the first few examples in the
paper and one from the running example on
extracting phone numbers.

@@

\subsection{Quotient Synthesis Algorithms}

% DPW:  From Quotient lenses paper

\newcommand{\wf}[1]{\ensuremath{#1\;\mathsf{wf}}}

% FOR Regular Expression names
\newcommand{\re}[1]{\ensuremath{\mathtt{#1}}}
\newcommand{\codefont}[1]{\ensuremath{\mathsf{#1}}}
\newcommand{\kw}[1]{\textcolor{dkblue}{\ensuremath{\mathsf{#1}}}}
\newcommand{\collapse}[2]{\ensuremath{\kw{collapse} \; #1 \mapsto #2}}
\newcommand{\squash}[3]{\ensuremath{\kw{squash} \; #1 \rightarrow #2\; \kw{using} \; #3}}
\newcommand{\perm}[2]{\ensuremath{\kw{perm}(#1)\; \kw{with}\; #2}}
\newcommand{\normalize}[3]{\ensuremath{\kw{normalize}(#1, #2, #3)}}
\newcommand{\eqrel}[1]{\ensuremath{\equiv_{#1}}}

\newcommand{\canonize}{\ensuremath{\kw{canonize}}}

\newcommand{\Name}{Optometrist\xspace}

\newcommand{\QRESize}{\textbf{QS}}
\newcommand{\canonizeAndSpecSize}{\textbf{BS}}
\newcommand{\LensAndSpecSize}{\textbf{NS}}

%\newcommand{\QOpt}{Optician_Q}
\newcommand{\QOpt}{QRE-enhanced Optician}
\newcommand{\OpticianRuntime}{\textbf{Optician}}
\newcommand{\QREOptician}{\textbf{Optician\textsubscript{Q}}}
\newcommand{\SystemOnOptician}{\textbf{QO}}
\newcommand{\SystemOnBenchmarks}{\textbf{QQ}}
\newcommand{\cd}[1]{\lstinline[backgroundcolor=\color{white}]$#1$}

We have implemented QREs and the quotient lens synthesis algorithm described
above as an extension to the Boomerang
interpreter~\cite{boomerang,quotientlenses}. We have extended the existing
Optician tool to synthesize QRE lenses from QREs. We will use 
``\QOpt'' to denote our extended version of Optician, and just plain ``Optician'' to denote
the pre-QRE version of Optician. The synthesis algorithm produces Boomerang lens
values, so Boomerang gives synthesized lenses the same first-class status as hand-written ones. To
evaluate the effectiveness of QREs, QRE lenses, and QRE lens synthesis, we
conducted experiments to answer the following questions:
%
\begin{itemize}
  \item {\bf Ease of use.} Does synthesizing QRE lenses from QREs permit an
  easier development process than writing lenses by hand?
  Does synthesizing QRE lenses from QREs permit an easier development
  process than manually writing canonizers and then synthesizing lenses between
  their canonized forms?  
  \item {\bf Performance.} Is the synthesis algorithm/implementation fast enough to be
  used as part of a standard development process?
\end{itemize}
%
All evaluations were performed on a 2.5 GHz Intel Core i7 processor with 16 GB
of 1600 MHz DDR3 running macOS High Sierra.

\paragraph*{Benchmark Suite Construction}

To evaluate our QRE implementation, we adapted 39 lens synthesis tasks
from the benchmark suite of the original Optician system. These
benchmarks are a combination of custom benchmarks, benchmarks derived
from FlashFill~\cite{flashfill}, and benchmarks derived from
Augeas~\cite{augeas2} (we also experimented using our QRE implementation to synthesize quotient lenses between XML, RDF and and JSON formats using data from the data.gov database; the data consisted of census statistics, demographic statistics, wage comparosion data, and crime index data).  In these 39 benchmarks, 10 of
the data formats had to be modified to work with the bijectivity
constraints that Optician required. For instance, when one
representation permits whitespace where the other does not, we
modified the original version of the benchmark to allow more whitespace,
thereby restoring bijectivity (but altering the data format). With the
new QRE support, we were able to remove these alterations.  This
experience alone suggests that QREs make the lens development process
more flexible.

\paragraph*{Ease of Use}

\begin{figure}[t]
\includegraphics{qfigs/asts.eps}
\caption{AST node measurements for each of the three approaches
on each of the 10 non-bijective benchmark problems.
Benchmarks are sorted in order of increasing complexity as measured by
the number of AST nodes in the source and target format descriptions. 
QRE Synthesis requires far fewer AST nodes than the other
two approaches.}
%%\caption{Count of benchmark programs definable using a given AST count. We find
%%that it takes far fewer AST nodes to define benchmark lenses using QRE
%%synthesis than with Optician without QRE synthesis or without synthesis.}
\label{fig:asts}
\end{figure}

To evaluate the impact of QRE lens synthsis on programmer effort,
we focus our attention on the 10 problems in the benchmark suite that
are not bijective and hence require non-trivial canonizers. 
(Optician already handles the other problems with minimal programmer
effort.)

We are interested in comparing three different approaches, which vary
in the amount of synthesis used. 
In the first approach, which we call \QRESize{} for QRE Synthesis, the programmer uses
QRE lens synthesis.  She must write QRE specifications of the source and
target formats and she may give examples.
In the second approach, which we call \canonizeAndSpecSize{} for
Bijective Synthesis, the
programmer uses bijective lens synthesis \`a la Optician.
She must write canonizers by hand, along with 
regular expressions to describe the external
representations of the source and target formats. (The internal
formats can be inferred from the canonizers.) She may also
provide examples to help in the synthesis of the bijective lens.
In the third approach, which we call \LensAndSpecSize{} for No Synthesis, the
programmer writes the lens between the source and target formats
entirely by hand, including the descriptions of the source and target
formats.

For each problem in the benchmark suite, we calculate the following
measures as proxies for the level of programmer effort when using each
the three approaches:

%
\begin{itemize}
  \item[\QRESize{}:] 
  The number of AST nodes in the QRE specifications for the source and
  target formats, including examples. 
  \item[\canonizeAndSpecSize{}:] 
  The sum of (1) the number of AST nodes in $W(q)$ for each QRE $q$ in the source and target
  formats, (2) the number of AST nodes in $\canonize(q)$ for each QRE $q$ with a
  non-trivial canonizer, and (3) the number of AST nodes in the
  examples.  We use (1) to estimate the burden of describing
  the external source and target formats and (2) to estimate the
  burden of writing the requisite canonizers
  by hand.  We count the nodes in the examples because they would be
  fed to the bijective synthesizer.  
  These counts are an approximation, as both $W(q)$ and $\canonize(q)$ are
  automatically generated from the corresponding QRE $q$, and it is
  possible that a human-written version might be smaller.
  \item[\LensAndSpecSize{}:] The sum of (1) the number of AST nodes in
  $W(q)$ for each QRE $q$ in the source and target formats and (2) the
  number of AST nodes in the synthesized QRE lens.  We use (1) to
  estimate the burdern of describing the source and target formats
  and (2) to estimate the burdern of writing the appropriate lens by
  hand. These counts are also approximations, as
  $W(q)$ and the synthesized lens may be larger than one written by hand.
\end{itemize}

Figure~\ref{fig:asts} shows each of these measures for the 10
non-bijective problems in the benchmark suite.  On
average (using a geometric mean), \canonizeAndSpecSize{} used~38.5\% more AST nodes
than \QRESize{}, requiring an average of~214 more AST nodes. On 
average, \LensAndSpecSize{} used~180\% more AST nodes than \QRESize{}, requiring an
average of~998 more AST nodes. These figures suggest that introducing QREs saves
programmers significant effort compared to both Optician and basic
Boomerang.

\paragraph*{Maintaining Competitive Performance}

\begin{figure}[t]
\centering
\begin{subfigure}[b]{.49\textwidth}
\centering
\includegraphics{qfigs/times_opt}
\caption{}
\label{subfig:lenssize}
\end{subfigure}
\begin{subfigure}[b]{.49\textwidth}
\includegraphics{qfigs/times_new.eps}
\caption{}
\label{subfig:examplesused}
\end{subfigure}
\caption{Runtimes measurements. In (a), we run Optician and \QOpt{}
  on the Optician benchmarks.  We find that there is only a negligible
  performance overhead incurred by using QREs. 
  In (b), we run \QOpt{} on the 10 Optician benchmarks
  previously edited to make them bijective, after removing those edits
  and then extending the synthesis specification to include QREs.
  (In other words, we restored them to their original state, added QREs,
  and then ran \QOpt{}). 
  We find that \QOpt{} is able to
  synthesize all quotient lenses in under 10 seconds, and typically finishes in
  under 5 seconds.}
\label{fig:times}
\end{figure}

To assess the performance of QRE synthesis, we are interested in two different
questions. First, how does the performance of \QOpt{} compare to the performance
of Optician on benchmarks that do not require QREs? The answer to this question
tells us how much overhead we have introduced by adopting the more general
mechanism. Figure~\ref{fig:times}(a) shows that \QOpt{} was able to synthesize
all of the Optician benchmarks at a speed competitive with the old version.
There is a small amount of additional overhead introduced by QREs in calculating
the $W$ and $K$ functions, resulting in a slight decrease in performance.

Second, how much time does it take for \QOpt{} to synthesize a QRE
lens when running on a non-bijective benchmark problem?  
Figure~\ref{fig:times}(b) shows the amount of time required to infer a
lens for each of the 10 benchmark programs with nontrivial quotients.  
We find that \QOpt{} is able to synthesize all quotient lenses in
under 10~seconds, and typically finishes in under 5~seconds.
